### 2020-02-04 thesis/paper.tex
more

@@

Alternatively, the user can also create a data-set himself from within the tool. It is possible to draw points onto the canvas when the ``SinglePoint'' option is selected in the top right, as in Figure \ref{fig:data-view}. For this, the user only needs to click wherever he wants points placed. It is important to note that if there are more than two dimensions all values for the not shown dimensions will not be set, which may cause a problem for the clustering algorithms later on. Another possibility is the use of the ELKIGenerator, which accepts a XML-style input for defining how to generate data. The definition for this format can be found in corresponding definitions file on the ELKI Github repository \cite{elkixml} or in the tools lib folder. This generator allows defining distributions on dimensions, the number of points for each generated distribution and many additional things like plane rotations. This can be quite useful when wanting to work with a quick sample data-set. In Figure \ref{fig:loaded-data} one can see how the data-view looks when data with cluster labels is shown.

@@

\section{Selection of Algorithms and Parameters}
bla

@@

\subsection{Clustering Workflows}
As many runs of clustering algorithms with different parameter ranges can be performed and repeatedly defining this workflow can be tedious, the tool allows exporting and importing these workflows from the cluster workflow window. For this the underlying clustering objects are serialized and saved into .CWF (Cluster Workflow Files). It is important to note, that these files do not include the parameter settings for random seed or the meta clustering. This is intentional as I do not consider them to be part of the workflow and they should be set by the user for every run, if the default values are not satisfactory.

\subsection{Pre-Clustered Data}
Due to the clustering algorithms possibly being time consuming, either because of their complexity or just sheer number of runs, the results of these algorithms can be saved from the meta-view. To do this, the clustering results are serialized and saved as .CRF (Clustering Result Files). To load those files one can define the different parameters of the meta clustering and then import the .CRF. This way, only the meta clustering is computed again. The meta clustering is not saved in this file such that the user can more easily try different settings for it, by just saving the result and running the meta clustering again. As this is usually much faster than the clustering algorithms that generate the base clusterings, using this may save a considerable amount of time.

@@

\subsubsection{Clustering Algorithms}
Clustering algorithms can be used in the workflow-view to generate base clusterings for meta clustering and visualization in the meta-view. The interface IClusterer is used as helper interface such that all clustering algorithms can use the same high level calls for compatibility. The actual implementations implement either IELKIClusterer if they use the ELKI database objects in their logic or ICustomClusterer if they use a data matrix. To further help not needing to re-implement generally needed methods the class AbstractClustering should also be extended by the actual implementation. With these class and interfaces the actual implementations can focus more on just the clustering logic. Additionally options panels should be created for each new algorithm allowing the user to set parameters or parameter ranges from the tool.

@@

\subsection{Meta-Clustering}
%info on OPTICS?
The algorithm used for meta clustering is OPTICS [CITATION]. One of the requirements for this methods is that the used distance measure must be metric, otherwise the algorithm can only produce an approximate clustering at best. The distance measure used by the algorithm can be changed and additional distance measures can also be added. For this, the following interface is of interest:

\begin{itemize}
  \item Clustering Distance Measure: IMetaDistanceMeasure
\end{itemize}
\subsubsection{Clustering Distance Measures}
The interface IMetaDistanceMeasure is used to define distance measures that can be used for OPTICS for meta clustering. These distance measures should represent how different two clustering solutions are from each other while needing to fulfill the requirements of a metric. For a distance measure, here indicated with $d(x,y)$ where $x$ and $y$ are objects, to be considered metric the following conditions must hold:
\begin{itemize}
  \item $d(x,y)\geq0$: distances must be non-negative
  \item $d(x,y)=0$ iff $x=y$: the distance between two objects is zero if and only if both objects are equal
  \item $d(x,y)=d(x,y)$: the distance must fulfill the symmetry condition
  \item $d(x,z)\leq d(x,y)+d(y,z)$: the triangle inequality must hold
\end{itemize}

@@

\subsection{Consensus Clustering}
Consensus Clustering can combine the results form different base clusterings to a common clustering. This can be done in the meta-view and different algorithms can be used for this step. When run, the algorithm computes a consensus result for all base clusterings that are in a common meta cluster in the OPTICS reachability plot and are not filtered out. This means that if the used defined three meta clusters with some cut-off value in the reachability plot, there will be three resulting consensus clusterings, one created from each group. The interface of interest for consensus functions is:

>>>>

\subsection{Plots and I/O}

@@

Alternatively, the user can also create a data-set himself from within the tool. It is possible to draw points onto the canvas when the ``SinglePoint'' option is selected in the top right, as in Figure \ref{fig:data-view}. For this, the user only needs to click wherever he wants points placed. It is important to note that if there are more than two dimensions all values for the not shown dimensions will not be set, which may cause a problem for the clustering algorithms later on. Another possibility is the use of the ``ELKIGenerator'', which accepts a XML-style input for defining how to generate data. The definition for this format can be found in corresponding definitions file on the ELKI Github repository \cite{elkixml} or in the tools lib folder. This generator allows defining distributions on dimensions, the number of points for each generated distribution and many additional things like plane rotations. This can be quite useful when wanting to work with a quick sample data-set. In Figure \ref{fig:loaded-data} one can see how the data-view looks when data with cluster labels is shown.

@@

\subsection{Modifying the data}

As the data may still contain unwanted dimensions after loading cleanup may need to be done. To reduce the dimensionality the simplest method is to select the ``Dim. Remover'' tool in the top right and define the index of the dimension to be removed. As more sophisticated methods PCA and t-SNE [CITATION] are also available. PCA calculates the principal components of the data and projects it onto the $k$ strongest ones leaving us with $k$ dimensions. $k$ can easily be defined in the settings of PCA by filling in the dimensions field, as seen in Figure \ref{fig:loaded-data}. t-SNE works similarly to PCA, has more options though and even a parallel implementation. For more information on those methods see Section [REFERENCE].

Other than reducing the dimensionality of the data the application allows to normalize the data, such that it can be more easily used for clustering. If there are different dimensions with strongly different meaning, unit and range this normalization step can benefit the accuracy of clustering greatly [CITATION?]. The available tools are ``Normalize'' and ``Standardize''. Normalize changes the value range for each dimension to the interval $[0,1]$ while preserving relative distances, the Standardize tool adapts these ranges such that the average is $0$ and the standard deviation is $1$.

With these to pre-processing steps done, the actual clustering of the data can be thought of next. To work on this the ``Clustering'' button opens the next major view of the tool.

\section{Selection of Algorithms and Parameters}
By clicking the ``Clustering'' button in the data-view the window managing clustering workflows, further on called workflow-view, opens. Here clustering algorithms, their parameters and the parameters for meta-clustering can be selected. The workflow-view can be seen in Figure [REFERENCE]

@@

\subsection{Clustering Workflows}
As many runs of clustering algorithms with different parameter ranges can be performed and repeatedly defining this workflow can be tedious, the tool allows exporting and importing these workflows from the cluster workflow window. For this the underlying clustering objects are serialized and saved into .CWF (Cluster Workflow Files). It is important to note, that these files do not include the parameter settings for random seed or the meta-clustering. This is intentional as I do not consider them to be part of the workflow and they should be set by the user for every run, if the default values are not satisfactory.

\subsection{Pre-Clustered Data}
Due to the clustering algorithms possibly being time consuming, either because of their complexity or just sheer number of runs, the results of these algorithms can be saved from the meta-view. To do this, the clustering results are serialized and saved as .CRF (Clustering Result Files). To load those files one can define the different parameters of the mata-clustering and then import the .CRF. This way, only the mata-clustering is computed again. The mata-clustering is not saved in this file such that the user can more easily try different settings for it, by just saving the result and running the mata-clustering again. As this is usually much faster than the clustering algorithms that generate the base clusterings, using this may save a considerable amount of time.

@@

\subsubsection{Clustering Algorithms}
Clustering algorithms can be used in the workflow-view to generate base clusterings for mata-clustering and visualization in the meta-view. The interface IClusterer is used as helper interface such that all clustering algorithms can use the same high level calls for compatibility. The actual implementations implement either IELKIClusterer if they use the ELKI database objects in their logic or ICustomClusterer if they use a data matrix. To further help not needing to re-implement generally needed methods the class AbstractClustering should also be extended by the actual implementation. With these class and interfaces the actual implementations can focus more on just the clustering logic. Additionally options panels should be created for each new algorithm allowing the user to set parameters or parameter ranges from the tool.

@@

\subsection{Meta-Clustering}
%info on OPTICS?
The algorithm used for mata-clustering is OPTICS [CITATION]. One of the requirements for this methods is that the used distance measure must be metric, otherwise the algorithm can only produce an approximate clustering at best. The distance measure used by the algorithm can be changed and additional distance measures can also be added. For this, the following interface is of interest:

\begin{itemize}
  \item Clustering Distance Measure: IMetaDistanceMeasure
\end{itemize}
\subsubsection{Clustering Distance Measures}
The interface IMetaDistanceMeasure is used to define distance measures that can be used for OPTICS for mata-clustering. These distance measures should represent how different two clustering solutions are from each other while needing to fulfill the requirements of a metric. For a distance measure, here indicated with $d(x,y)$ where $x$ and $y$ are objects, to be considered metric the following conditions must hold:
\begin{itemize}
  \item $d(x,y)\geq0$: distances must be non-negative
  \item $d(x,y)=0$ iff $x=y$: the distance between two objects is zero if and only if both objects are equal
  \item $d(x,y)=d(x,y)$: the distance must fulfill the symmetry condition
  \item $d(x,z)\leq d(x,y)+d(y,z)$: the triangle inequality must hold
\end{itemize}

@@

\subsection{Consensus Clustering}
Consensus Clustering can combine the results form different base clusterings to a common clustering. This can be done in the meta-view and different algorithms can be used for this step. When run, the algorithm computes a consensus result for all base clusterings that are in a common mata-cluster in the OPTICS reachability plot and are not filtered out. This means that if the used defined three mata-clusters with some cut-off value in the reachability plot, there will be three resulting consensus clusterings, one created from each group. The interface of interest for consensus functions is:
